---
title: "Machine learning Project Assignment"
author: "Samer Mounir"
date: "June 19, 2015"
output: html_document
---
Document is produced under R version 3.2.0 & Windows 7
Document purpose is to solve Coursera Machine learning Project Assignment

###1. we need to prepare load our data libraries
```{r datalib}

library (caret)
library (rattle)
library (randomForest)
# enable multi-core processing, Thanks to Coursera Discussion forum
library(doParallel)
cl <- makeCluster(detectCores())
registerDoParallel(cl)
```
###2. Download the data from the specified URL , then start reading the data and exploring it, and cleanse it

```{r Reading, cache=TRUE}
## Set the data seed  for reproducability
set.seed(1234)
# Setting the Working directory
setwd ("C://Personal//Coursera//RDir//8MachineLearning")
# Reading the Training & Testing files and replacing the NA, #DIV/0! & empty with NA
training <- read.csv("pml-training.csv",na.strings=c("NA","#DIV/0!", ""))
testing<-read.csv("pml-testing.csv",na.strings=c("NA","#DIV/0!", ""))

# Removing the columns that are filled with NA from the training and testing data sets
ctraining<-training[,colSums(is.na(training)) == 0]
ctesting<-testing[,colSums(is.na(testing)) == 0]
# Exploring the dataset
dim(ctraining)

# Removing the first 7 coulmns as they are not relevant with our work
ctraining   <-ctraining[,-c(1:7)]
ctesting <-ctesting[,-c(1:7)]

```
The final training set contains 53 Variables with 19622 objects

###3. Split the Training data set into a training and crossvalidation dat set
We are going to split the data between training & cross validation in the ratio of 70 to 30
```{r datasplit}
intrain <- createDataPartition(y=ctraining$classe,p=0.7,list=FALSE)
trainpart <- ctraining[intrain,]
crosspart <- ctraining [-intrain,]
```
###4. Testing different prediction model
Hence the required prediction is certain values I have decided to test the Random forest model


``` {r decisoinTree,cache=TRUE}
# Train  the model
modelfit <- randomForest(classe ~. , data=trainpart, method="class")
# Cross Check the model
prediction <- predict(modelfit, crosspart, type = "class")
# Testing the model on the crosspart

confusionMatrix(prediction, crosspart$classe)
```
The results of the random forest model is pretty high
Accuracy : 0.9966  95% CI : (0.9948, 0.9979) 
which gives  confidence in this model

###5 Acquiring the results for the testing data 
After this we need to apply the model on the 20 records of the test data and acquire the results

```{r Predictingresults}
predictfinal <- predict(modelfit, ctesting,type="class")
predictfinal
```

* The accuracy of the model is 0.996 or 99.6%
* The expected out-of-sample error is estimated at 0.005, or 0.5% as it is calculated as 1 - accuracy for predictions made against the cross-validation set.
* Our Test data set comprises 20 cases. With an accuracy above 99% on our cross-validation data, we can expect that very few, or none, of the test samples will be missclassified.

###6 Creating the files to be uploaded into coursera website

```{r files}
pml_write_files = function(x){
      n = length(x)
      for(i in 1:n){
            filename = paste0("problem_id_",i,".txt")
            write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
      }
}
#pml_write_files(predictfinal)
```
